\textbf{COMP0086 Summative Assignment}

### 2022-11-14 submission.tex
remove extra R

\item[(b)] We can now reformulate the optimisation as:
$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$$
Because $\| \textbf{x}\|=1$ (i.e. choosing $\textbf{w} \in \mathbb{R}^n$ and $\textbf{x}=\frac{\textbf{w}}{\|\textbf{w}\|}$ to reformulate the problem over the unit sphere), we can equivalently write:
$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \textbf{x}^T\textbf{Ax}}$$
Thus, showing $R_A(\textbf{x}) \leq \lambda_1$ will be equivalent to showing $\textbf{x}^T\textbf{Ax}} \leq \lambda_1$ for $\|\textbf{x}\|=1$.
We know that for all $\textbf{x} \in \mathbb{R}^n$:
$$\textbf{x} = \sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i$$
so we can write:
$$\textbf{x}^T\texRtbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\textbf{A}\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i\bigg)$$
Given that $\xi_i$R are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\bigg(\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x}) \xi_i\bigg)$$
Given that the eigenvectors $\xi_i$ form an orthonormal basis, we know that $\xi_i^T\xi_j = 0$ when $i\neq j$ and $\xi_i^T\xi_j = 1$ when $i=j$, so:
$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$$
From our above reformulation with the unit sphere, we know that $\|\textbf{x}\|^2 = 1$ so $\|\textbf{x}\|^2 = \sum_{i=1}^{n} x_i^2 = \sum_{j=i}^{n} (\xiij \textbf{x})^2 = 1$. Thus the quantity $\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$ is a weighted average of $\lambda_i$'s with weights $(\xi_i^T \textbf{x})^2$, which is always less than or equal to the largest $\lambda_i$ value so:
$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \lambda_1$$
where $\lambda_1$ is the largest eigenvalue of eigenvalues $\lambda_i$. Therefore, $R_A(\textbf{x}) \leq \lambda_1$ as required.

>>>>

\item[(b)] We can now reformulate the optimisation as:
$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$$
Because $\| \textbf{x}\|=1$ (i.e. choosing $\textbf{w} \in \mathbb{R}^n$ and $\textbf{x}=\frac{\textbf{w}}{\|\textbf{w}\|}$ to reformulate the problem over the unit sphere), we can equivalently write:
$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \textbf{x}^T\textbf{Ax}}$$
Thus, showing $R_A(\textbf{x}) \leq \lambda_1$ will be equivalent to showing $\textbf{x}^T\textbf{Ax}} \leq \lambda_1$ for $\|\textbf{x}\|=1$.
We know that for all $\textbf{x} \in \mathbb{R}^n$:
$$\textbf{x} = \sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i$$
so we can write:
$$\textbf{x}^T\texRtbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\textbf{A}\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i\bigg)$$
Given that $\xi_i$ are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\bigg(\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x}) \xi_i\bigg)$$
Given that the eigenvectors $\xi_i$ form an orthonormal basis, we know that $\xi_i^T\xi_j = 0$ when $i\neq j$ and $\xi_i^T\xi_j = 1$ when $i=j$, so:
$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$$
From our above reformulation with the unit sphere, we know that $\|\textbf{x}\|^2 = 1$ so $\|\textbf{x}\|^2 = \sum_{i=1}^{n} x_i^2 = \sum_{j=i}^{n} (\xiij \textbf{x})^2 = 1$. Thus the quantity $\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$ is a weighted average of $\lambda_i$'s with weights $(\xi_i^T \textbf{x})^2$, which is always less than or equal to the largest $\lambda_i$ value so:
$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \lambda_1$$
where $\lambda_1$ is the largest eigenvalue of eigenvalues $\lambda_i$. Therefore, $R_A(\textbf{x}) \leq \lambda_1$ as required.
